Capstone Movielens Report

### 2025-05-07 reports/capstone-movielens-report.draft4.tex
Draft report updated.

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-15-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-19-1.pdf}

@@

\subsubsection{Defining Logging and Mesuring helper
functions}\label{defining-logging-and-mesuring-helper-functions}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{open\_logfile }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(file\_name)\{}
\NormalTok{  log\_file\_name }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(}\FunctionTok{Sys.time}\NormalTok{()) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{str\_replace\_all}\NormalTok{(}\StringTok{\textquotesingle{}:\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}\_\textquotesingle{}}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{str\_replace}\NormalTok{(}\StringTok{\textquotesingle{} \textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}T\textquotesingle{}}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{str\_c}\NormalTok{(file\_name)}

@@

\NormalTok{print\_log }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg)\{}
  \FunctionTok{print}\NormalTok{(}\FunctionTok{str\_glue}\NormalTok{(msg))}
\NormalTok{\}}
\NormalTok{put\_log }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg)\{}
  \FunctionTok{put}\NormalTok{(}\FunctionTok{str\_glue}\NormalTok{(msg))}
\NormalTok{\}}
\NormalTok{put\_log1 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg1)\{}
\NormalTok{  msg }\OtherTok{\textless{}{-}} \FunctionTok{str\_replace\_all}\NormalTok{(msg\_template, }\StringTok{"\%1"}\NormalTok{, }\FunctionTok{as.character}\NormalTok{(arg1))}
  \FunctionTok{put}\NormalTok{(}\FunctionTok{str\_glue}\NormalTok{(msg))}
\NormalTok{\}}
\NormalTok{put\_log2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg1, arg2)\{}
\NormalTok{  msg }\OtherTok{\textless{}{-}}\NormalTok{ msg\_template }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{str\_replace\_all}\NormalTok{(}\StringTok{"\%1"}\NormalTok{, }\FunctionTok{as.character}\NormalTok{(arg1)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{str\_replace\_all}\NormalTok{(}\StringTok{"\%2"}\NormalTok{, }\FunctionTok{as.character}\NormalTok{(arg2)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{str\_glue}\NormalTok{()}

  \FunctionTok{put}\NormalTok{(msg)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  mu }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(edx}\SpecialCharTok{$}\NormalTok{rating)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{print}\NormalTok{(mu)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 3.47184
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu.MSEs }\OtherTok{\textless{}{-}} \FunctionTok{naive\_model\_MSEs}\NormalTok{(mu)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
  \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{fold\_No =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{5}\NormalTok{, }\AttributeTok{MSE =}\NormalTok{ mu.MSEs) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{data.plot}\NormalTok{(}\AttributeTok{title =} \StringTok{"MSE resuls of the 5{-}fold CV method applied to the Overall Mean Rating Model"}\NormalTok{,}
                \AttributeTok{xname =} \StringTok{"fold\_No"}\NormalTok{, }
                \AttributeTok{yname =} \StringTok{"MSE"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-40-1.pdf}

@@

\begin{verbatim}
## [1] 1.061191
\end{verbatim}

@@

\NormalTok{  deviation.RMSE }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(deviation, }\ControlFlowTok{function}\NormalTok{(delta)\{}
    \FunctionTok{naive\_model\_RMSE}\NormalTok{(mu }\SpecialCharTok{+}\NormalTok{ delta)}
\NormalTok{  \})}
  \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"RMSE values have been computed for \%1 deviations from the Overall Mean Rating."}\NormalTok{,}
           \FunctionTok{length}\NormalTok{(deviation))}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
## Warning in is.na(xlabel): is.na() applied to non-(list or vector) of type 'expression'
\end{verbatim}

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-44-1.pdf}

@@

\FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Minimum RMSE is achieved when the deviation from the mean is: \%1"}\NormalTok{,}
\NormalTok{         which\_min\_deviation)}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
## Log is not open.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Is the previously computed RMSE the best for the current model: \%1"}\NormalTok{,}
\NormalTok{         mu.RMSE }\SpecialCharTok{==}\NormalTok{ min\_rmse)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Is the previously computed RMSE the best for the current model: FALSE
\end{verbatim}

\begin{verbatim}
## Log is not open.
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-57-1.pdf}

>>>>

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-22-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-25-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-26-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-29-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-30-1.pdf}

@@

\subsubsection{Defining Logging and Time Measuring Helper
Functions}\label{defining-logging-and-time-measuring-helper-functions}

@@

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# Logging Helper functions {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\NormalTok{open\_logfile }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(file\_name)\{}
\NormalTok{  log\_file\_name }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(}\FunctionTok{Sys.time}\NormalTok{()) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{str\_replace\_all}\NormalTok{(}\StringTok{\textquotesingle{}:\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}\_\textquotesingle{}}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{str\_replace}\NormalTok{(}\StringTok{\textquotesingle{} \textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}T\textquotesingle{}}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{str\_c}\NormalTok{(file\_name)}

@@

\NormalTok{msg.set\_arg }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg, }\AttributeTok{arg.name =} \StringTok{"\%1"}\NormalTok{) \{}
\NormalTok{  msg\_template }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{str\_replace\_all}\NormalTok{(arg.name, }\FunctionTok{as.character}\NormalTok{(arg))}
\NormalTok{\}}
\NormalTok{msg.glue }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg, }\AttributeTok{arg.name =} \StringTok{"\%1"}\NormalTok{)\{}
\NormalTok{  msg\_template }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{msg.set\_arg}\NormalTok{(arg, arg.name) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{str\_glue}\NormalTok{()}
\NormalTok{\}}

\NormalTok{print\_log }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg)\{}
  \FunctionTok{print}\NormalTok{(}\FunctionTok{str\_glue}\NormalTok{(msg))}
\NormalTok{\}}
\NormalTok{put\_log }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg)\{}
  \FunctionTok{put}\NormalTok{(}\FunctionTok{str\_glue}\NormalTok{(msg))}
\NormalTok{\}}

\NormalTok{get\_log1 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg1) \{}
  \FunctionTok{str\_glue}\NormalTok{(}\FunctionTok{str\_replace\_all}\NormalTok{(msg\_template, }\StringTok{"\%1"}\NormalTok{, }\FunctionTok{as.character}\NormalTok{(arg1)))}
\NormalTok{\}}
\NormalTok{print\_log1 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg1)\{}
  \FunctionTok{print}\NormalTok{(}\FunctionTok{get\_log1}\NormalTok{(msg\_template, arg1))}
\NormalTok{\}}
\NormalTok{put\_log1 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg1)\{}
  \FunctionTok{put}\NormalTok{(}\FunctionTok{get\_log1}\NormalTok{(msg\_template, arg1))}
\NormalTok{\}}

\NormalTok{get\_log2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg1, arg2) \{}
\NormalTok{  msg\_template }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{str\_replace\_all}\NormalTok{(}\StringTok{"\%1"}\NormalTok{, }\FunctionTok{as.character}\NormalTok{(arg1)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{str\_replace\_all}\NormalTok{(}\StringTok{"\%2"}\NormalTok{, }\FunctionTok{as.character}\NormalTok{(arg2)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{str\_glue}\NormalTok{()}
\NormalTok{\}}
\NormalTok{print\_log2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg1, arg2)\{}
  \FunctionTok{print}\NormalTok{(}\FunctionTok{get\_log1}\NormalTok{(msg\_template, arg1, arg2))}
\NormalTok{\}}
\NormalTok{put\_log2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(msg\_template, arg1, arg2)\{}
  \FunctionTok{put}\NormalTok{(}\FunctionTok{get\_log1}\NormalTok{(msg\_template, arg1, arg2))}
\NormalTok{\}}

\CommentTok{\# ...}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(edx}\SpecialCharTok{$}\NormalTok{rating)}
\FunctionTok{print}\NormalTok{(mu)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 3.512465
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu.MSEs }\OtherTok{\textless{}{-}} \FunctionTok{naive\_model\_MSEs}\NormalTok{(mu)}
\FunctionTok{data.frame}\NormalTok{(}\AttributeTok{fold\_No =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{5}\NormalTok{, }\AttributeTok{MSE =}\NormalTok{ mu.MSEs) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{data.plot}\NormalTok{(}\AttributeTok{title =} \StringTok{"MSE resuls of the 5{-}fold CV method applied to the Overall Mean Rating Model"}\NormalTok{,}
              \AttributeTok{xname =} \StringTok{"fold\_No"}\NormalTok{, }
              \AttributeTok{yname =} \StringTok{"MSE"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-49-1.pdf}

@@

\begin{verbatim}
## [1] 1.060346
\end{verbatim}

@@

\NormalTok{  deviation.RMSE }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(deviation, }\ControlFlowTok{function}\NormalTok{(delta)\{}
    \FunctionTok{naive\_model\_RMSE}\NormalTok{(mu }\SpecialCharTok{+}\NormalTok{ delta)}
\NormalTok{  \})}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-53-1.pdf}

@@

\FunctionTok{print\_log1}\NormalTok{(}\StringTok{"Minimum RMSE is achieved when the deviation from the mean is: \%1"}\NormalTok{,}
\NormalTok{         which\_min\_deviation)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{print\_log1}\NormalTok{(}\StringTok{"Is the previously computed RMSE the best for the current model: \%1"}\NormalTok{,}
\NormalTok{         mu.RMSE }\SpecialCharTok{==}\NormalTok{ min\_rmse)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Is the previously computed RMSE the best for the current model: TRUE
\end{verbatim}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-54-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-66-1.pdf}
